---
title: "NBER Submissions"
author: "Ashwin Malshe"
date: "10/1/2021"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)

papers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/papers.csv')
authors <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/authors.csv')
programs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/programs.csv')
paper_authors <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/paper_authors.csv')
paper_programs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/paper_programs.csv')

theme_set(ggthemes::theme_fivethirtyeight())

```

## Visualization of monthly NBER paper submission

Anecdotally I had heard that many economists were submitting tons of papers to NBER during the Covid-19 pandemic. However, it was not clear to me whether this was a trend or a flash in the pan. The following visualization show that the spike in 2020 was primarily in one month. 

```{r}
papers %>% 
  count(year, month) %>% 
  mutate(date = as.Date(paste(year, month, "01", sep = "-"))) %>% 
  ggplot(aes(x = date, y = n)) +
  geom_line(size = 0.3) 
  
```

Which month saw the spike? This table shows that it was in May 2020. Top 6 months with highest submissions are in 2020.

```{r}
papers %>% 
  count(year, month, sort = TRUE) %>% 
  head(10)
```

Let's annotate the earlier plot with the spike in may 2020.

```{r}
papers %>% 
  count(year, month) %>% 
  mutate(date = as.Date(paste(year, month, "01", sep = "-"))) %>% 
  ggplot(aes(x = date, y = n)) +
  geom_line(size = 0.3) +
  annotate(geom = "text", x = as.Date("2015-01-01"), y = 223, 
           label = "May 2020 had \n223 submissions", color = "#b131a2", fontface = "italic")
```

## Sub-areas of Economics submitting in May 2020

Next, let's visualize the top subcategories submitting in May 2020. I expected these to be from Heath Care and Heath Economics. But is that the case?

Create a merged data set with `papers` and `programs`
```{r}

papers_merged <- papers %>% 
  left_join(paper_programs, by = "paper") %>% 
  left_join(programs, by = "program")
```


```{r}
top_10_may_2020 <- papers_merged %>% 
  filter(year == 2020 & month == 5) %>% 
  count(program_desc, sort = TRUE) %>% 
  head(10)
```

```{r}
top_10_may_2020
```



```{r}
top_10_may_2020 %>% 
  ggplot(aes(x = reorder(program_desc, n), y = n )) +
  geom_col() +
  coord_flip() 
  
```

Let's fill the first 3 bars with a different color. Also, using `str_swap` function from `sringr` package (bundled with `tidyvere` so you don't have to install it separately) wrap the axis labels.


```{r}
top_10_may_2020 %>% 
  mutate(top3 = ifelse(n >= 56, 1, 0)) %>% 
  ggplot(aes(x = reorder(program_desc, n), y = n )) +
  geom_col(aes(fill = as.factor(top3)), show.legend = FALSE) +
  coord_flip() +
  scale_fill_manual(values = c("#f78ae0", "#b131a2")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 25))
  
```

How does the distribution of the top 10 most submitted program descriptions look like? First, get the top 10 most popular programs of all times.

```{r}
top_10 <-  papers_merged %>% 
  count(program_desc, sort = TRUE) %>%  
  head(10)
```

We can compare individual distributions using `facet_wrap`. But there is a better way to compare these distributions over time using `ggridges` package.


```{r}

papers_merged %>% 
  inner_join(top_10, by = "program_desc") %>% 
  ggplot(aes(x = year, y = program_desc, fill = program)) +
  geom_density_ridges(scale = 4, alpha = 0.5, size = 0.3) + 
  scale_fill_cyclical(values = c("#6638f0", "#5cc9f5")) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 25))
  
  
```

Although almost all the top-10 fields experienced increases in 2020 submissions, International Trade and Finance related submissions were flat. Corporate Finance submissions saw a sharper bump compared to Asset Pricing, which both belong to Finance.


